### 2019-12-05 boek.tex
~

% use Sonny chapter style
\usepackage[Sonny]{fncychap} %Sonny

% set font after chapter sonny style
\usepackage{fontspec}
\setmainfont [Path = fonts/,
UprightFont = *-300,
ItalicFont = *-300-Italic,
BoldFont = *-700,
BoldItalicFont = *-700-Italic
]{MuseoSans}

@@

\hypersetup{
	colorlinks = true,
%	linkcolor = {blue},
  urlcolor = {blue},
	citecolor = {med-gray}
}

@@

%\renderalltrue

@@

\ifrenderall
\newpage
\graphicspath{{./ch2_dyngen/}}
\include{ch2_dyngen/content}
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%     CH3: DYNBENCHMARK     %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{dynbenchmark: A comparison of single-cell trajectory inference methods} 
\chaptermark{dynbenchmark: A comparison of single-cell trajectory inference methods.}
\label{chap:dynbenchmark}

{\large \textbf{Abstract}}
\begin{quote}
	\textbf{Purpose:} Trajectory inference approaches analyse genome-wide omics data from thousands of single cells and computationally infer the order of these cells along developmental trajectories. Although more than 75 trajectory inference tools have already been developed, it is challenging to compare their performance because the input they require and output models they produce vary substantially. \\
	\textbf{Results:} Here, we benchmark 45 of these methods on 110 real and 229 synthetic datasets for cellular ordering, topology, scalability and usability. Our results highlight the complementarity of existing tools, and that the choice of method should depend mostly on the dataset dimensions and trajectory topology. We develop a set of guidelines to help users select the best method for their dataset. \\
	\textbf{Conclusion:}  Our freely available data and evaluation pipeline will aid in the development of improved tools designed to analyse increasingly large and complex single-cell datasets.
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%       CH5: SCORPIUS       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{SCORPIUS: Fast, accurate, and robust single-cell pseudotime} 
\chaptermark{SCORPIUS: Fast, accurate, and robust single-cell pseudotime.} % short title
\label{chap:scorpius}

@@

{\large \textbf{Abstract}}
\begin{quote}
	\textbf{Purpose:} Graphlets are small network patterns that can be counted in order to characterise the structure of a network (topology). As part of a topology optimisation process, graphlet counts are used to iteratively modify a network in order to achieve certain topological properties.
	Up until now, however, graphlets were not suited as a metric for performing topology optimisation; when millions of minor changes are made to the network structure it becomes computationally intractable to recalculate all the graphlet counts for each of the edge modifications. \\
	\textbf{Results:} IncGraph calculates the differences in graphlet counts with respect to the network in its previous state.
	IncGraph reduces the execution time by several orders of magnitude when compared to recounting the graphlet counts from scratch at every iteration.
	We demonstrate the usefulness of this approach optimising gene regulatory networks using a graphlet-based fitness score. \\
	\textbf{Conclusion:} IncGraph is able to quickly quantify the topological impact of small changes to a network, which opens novel research opportunities to study changes in topologies in evolving or online networks, or develop graphlet-based criteria for topology optimisation.
\end{quote}

@@

\vfill
{\large \textbf{Publication status}}
\begin{quote}
  Published in Genome Biology 20, 125 (2019). \doi{10.1186/s13059-019-1738-8}. \\
	Weber LM, Saelens W, \textbf{Cannoodt R}, Soneson C, Hapfelmeier A, Gardner PP, Boulesteix AL, Saeys Y, and Robinson MD 
\end{quote}

>>>>

\usepackage{titlesec}
\newcommand{\hsp}{\hspace{20pt}}	
\titleformat{\chapter}[hang]{\LARGE\bfseries}{\thechapter\hsp\textcolor{light-gray}{|}\hsp}{0pt}{\LARGE\bfseries}

% set fonts
\usepackage{fontspec}
\setmainfont [Path = fonts/,
UprightFont = *-300,
ItalicFont = *-300-Italic,
BoldFont = *-700,
BoldItalicFont = *-700-Italic
]{MuseoSans}

@@

\hypersetup{
	colorlinks = true,
	linkcolor = {black},
  urlcolor = {black},
	citecolor = {med-gray}
}

@@

\renderalltrue

@@

\ifrenderall
\newpage
\graphicspath{{./ch2_dyngen/}}
\include{ch2_dyngen/content}
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%     CH3: DYNBENCHMARK     %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{dynbenchmark: A comparison of single-cell trajectory inference methods} 
\chaptermark{dynbenchmark: A comparison of single-cell trajectory inference methods.}
\label{chap:dynbenchmark}
\vspace*{-3em}
{\large \textbf{Abstract}}
\begin{quote}
	\textbf{Purpose:} Trajectory inference approaches analyse genome-wide omics data from thousands of single cells and computationally infer the order of these cells along developmental trajectories. Although more than 75 trajectory inference tools have already been developed, it is challenging to compare their performance because the input they require and output models they produce vary substantially. \\
	\textbf{Results:} Here, we benchmark 45 of these methods on 110 real and 229 synthetic datasets for cellular ordering, topology, scalability and usability. Our results highlight the complementarity of existing tools, and that the choice of method should depend mostly on the dataset dimensions and trajectory topology. We develop a set of guidelines to help users select the best method for their dataset. \\
	\textbf{Conclusion:}  Our freely available data and evaluation pipeline will aid in the development of improved tools designed to analyse increasingly large and complex single-cell datasets.
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%       CH5: SCORPIUS       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{SCORPIUS: Fast, accurate, and robust single-cell pseudotime} 
\chaptermark{SCORPIUS: Fast, accurate, and robust single-cell pseudotime.} % short title
\label{chap:scorpius}
\vspace*{-4em}

@@

{\large \textbf{Abstract}}
\begin{quote}
	\textbf{Purpose:} Graphlets are small network patterns that can be counted in order to characterise the structure of a network (topology). Counting graphlet can used to iteratively optimise a network in order to achieve certain topological properties.
	Up until now, however, graphlets were not suited as a metric for performing topology optimisation; when millions of minor changes are made to the network structure it becomes computationally intractable to recalculate all the graphlet counts for each of the edge modifications. \\
	\textbf{Results:} IncGraph calculates the differences in graphlet counts with respect to the network in its previous state.
	IncGraph reduces the execution time by several orders of magnitude when compared to recounting the graphlet counts from scratch at every iteration.
	We demonstrate the usefulness of this approach optimising gene regulatory networks using a graphlet-based fitness score. \\
	\textbf{Conclusion:} IncGraph is able to quickly quantify the topological impact of small changes to a network, which opens novel research opportunities to study changes in topologies in evolving or online networks, or develop graphlet-based criteria for topology optimisation.
\end{quote}

@@

\vfill
{\large \textbf{Publication status}}
\begin{quote}
  Published in Genome Biology 20 (2019). \doi{10.1186/s13059-019-1738-8}. \\
	Weber LM, Saelens W, \textbf{Cannoodt R}, Soneson C, Hapfelmeier A, Gardner PP, Boulesteix AL, Saeys Y, and Robinson MD 
\end{quote}

### 2019-12-05 ch0_frontmatter.tex
added github & orcid
~

@@

	{\LARGE \textsf{Robrecht Cannoodt}}

@@

\chapter*{Summary}
\vspace{-2em}

Recent developments in the \textit{single-cell omics} technologies triggered a major milestone in the field of (cell)biology. Current techniques can be used to profile both the genome and the transcriptome of individual cells, leading to important insights into cellular heterogeneity and dynamic cell processes. This evolution has led, among other things, to the Human Cell Atlas initiative, an international collaboration project with the aim of mapping all human cell types in terms of its biomolecular contents.

@@

\chapter*{Samenvatting}
\vspace{-8em}

>>>>

\newcommand{\orcid}[1]{	\includegraphics[width=3mm]{fig/icons/orcid.png}\;\href{https://orcid.org/#1}{{https://orcid.org/}{\color{black}#1}}}
\newcommand{\github}[1]{	\includegraphics[width=3mm]{fig/icons/github.png}\;\href{https://github.com/#1}{{https://github.com/}{\color{black}#1}}}

@@

	{\LARGE \textsf{Robrecht Cannoodt}} \\
	\orcid{0000-0003-3641-729X} \\
	\github{rcannood}

@@

\newpage
\ 
\vfill
{\bfseries\LARGE Summary}
\vfill
Recent developments in the \textit{single-cell omics} technologies triggered a major milestone in the field of (cell)biology. Current techniques can be used to profile both the genome and the transcriptome of individual cells, leading to important insights into cellular heterogeneity and dynamic cell processes. This evolution has led, among other things, to the Human Cell Atlas initiative, an international collaboration project with the aim of mapping all human cell types in terms of its biomolecular contents.

@@

\newpage
%\ 
%\vfill
{\bfseries\LARGE Samenvatting}
%\vfill

### 2019-12-05 chX_backmatter.tex
~

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{Curriculum Vitae}
\section{Personalia}
\begin{tabular}{ll}
	Name & Robrecht Cannoodt \\
	Date of Birth & January 25th, 1990 \\
	Place of birth & Ghent \\
	Nationality & Belgian \\
	Address & Kerkstraat 75, 9070 Destelbergen \\
	Email & \href{mailto:robrecht.cannoodt@gmail.com}{robrecht.cannoodt@gmail.com}
\end{tabular}

\section{Professional Experience}
\textit{Ph.D. student in Computer Science} \\
\hspace*{2 ex} Ghent University, Belgium, September 2013 -- now \\
\textit{Laboratory assistant (summer job)} \\
\hspace*{2 ex} Bloedtransfusiecentrum Gent, Belgium, July (2009 -- 2012)

@@

\begin{table}[ht!]
	\caption{\textbf{Contibutions to open-source software.} Following abbreviations denote the relation with respect to the package: \textit{aut} Author, \textit{ctb} Contributor. Yearly download statistics are based on the number of downloads between 2019-10-01 and 2019-10-21. CRAN download statistics are retrieved from the Rstudio CRAN mirror only; other CRAN mirrors do not track download statistics. In addition, many of the dynverse packages have only recently been published on CRAN. For Github repositories, no download statistics could be retrieved. } \label{tab:packages}

	\centering\fontsize{9}{11}\selectfont
	\begin{tabularx}{\linewidth}{|p{2cm}llp{1.5cm}X|}
		\hline
		Name & Role & Host & Downloads per year & Description \\ \hline\hline
		\cranpkg{babelwhale} & aut & CRAN & 9800 & Interacting with Docker and Singularity containers \\
		\cranpkg{diffusionMap} & aut & CRAN & 27'850 & Implements diffusion map method of data parameterization \\
		\githubpkg{dynverse}{dynbenchmark} & aut & Github & \notavailable & Pipeline for benchmarking trajectory inference methods \\
		\cranpkg{dyndimred} & aut & CRAN & 4708 & Applying dimensionality reduction methods \\
		\githubpkg{dynverse}{dyneval} & aut & Github & \notavailable & Evaluating trajectory inference methods \\
		\githubpkg{dynverse}{dynfeature} & aut & Github & \notavailable & Calculating feature importance scores from trajectories \\
		\githubpkg{dynverse}{dyngen} & aut & Github & \notavailable & Simulating single-cell data using gene regulatory networks \\
		\githubpkg{dynverse}{dynguidelines} & aut & Github & \notavailable & User guidelines for trajectory inference \\
		\githubpkg{dynverse}{dynmethods} & aut & Github & \notavailable & A collection of wrappers for trajectory inference methods \\
		\githubpkg{dynverse}{dyno} & aut & Github & \notavailable & A pipeline for inferring, visualising and interpreting trajectories \\
		\cranpkg{dynparam} & aut & CRAN & 2810 & Creating meta-information for parameters \\
		\githubpkg{dynverse}{dynplot} & aut & Github & \notavailable & A simple visualisation library for trajectories \\
		\githubpkg{dynverse}{dynplot2} & aut & Github & \notavailable & A fully customisable visualisation library for trajectories \\
		\githubpkg{dynverse}{dyntoy} & aut & Github & \notavailable & Generating simple toy data of cellular differentiation \\
		\cranpkg{dynutils} & aut & CRAN & 32'668 & Common functionality for the dynverse packages \\
		\cranpkg{dynwrap} & aut & CRAN & 4636 & A common format for trajectories \\
		\cranpkg{GillespieSSA} & aut & CRAN & 7118 & Gillespie's Stochastic Simulation Algorithm (SSA) \\
		\cranpkg{GillespieSSA2} & aut & CRAN & 4950 & Gillespie's Stochastic Simulation Algorithm for Impatient People \\
		\githubpkg{dynverse}{gng} & aut & Github & \notavailable & Growing Neural Gas implemented in Rcpp \\
		\cranpkg{incgraph} & aut & CRAN & 3565 & Incremental graphlet counting for network optimisation \\
		\cranpkg{lmds} & aut & CRAN & 23'' & Landmark Multi-Dimensional Scaling \\
		\cranpkg{princurve} & aut & CRAN & 29'100 & Fits a principal curve in arbitrary dimension \\
		\cranpkg{proxyC} & aut & CRAN & 117'480 & Computes proximity in large sparse matrices \\
		\cranpkg{qsub} & aut & CRAN & 3585 & Running commands remotely on gridengine clusters \\
		\cranpkg{SCORPIUS} & aut & CRAN & 4520 & Inferring developmental chronologies from single-cell RNA sequencing data \\ \hline\hline

		\cranpkg{badger} & ctb & CRAN & 6240 & Query information and generate badge for using in README \\
		\biocpkgl{ClusterSignificance}{Clus\-ter\-Sig\-nif\-i\-cance} &  & Bioc & 935 & Assess if class clusters in dimensionality reduced data representations have a separation different from permuted data \\
		\cranpkg{devtools} & ctb & CRAN & 5'918'700 & Tools to make developing R packages easier \\
		\githubpkg{soedinglab}{merlot} & ctb & Github & \notavailable & Reconstructing lineage-tree topologies from scRNA-seq data \\
		\cranpkg{mlr} & ctb & CRAN & 176'330 & Machine Learning in R \\
		\biocpkg{monocle} & ctb & Bioc & 34'360 & Clustering, differential expression, and trajectory analysis for single-cell RNA-Seq \\
		\cranpkg{ParamHelpers} & ctb & CRAN & 150'775 & Helpers for Parameters in Black-Box Optimization, Tuning and Machine Learning \\
		\githubpkg{kieranrcampbell}{pseudogp} & ctb & Github & \notavailable & Probabilistic pseudotime for single-cell RNA-seq \\
		\cranpkg{ranger} & ctb & CRAN & 413'641 & A Fast Implementation of Random Forests \\
		\cranpkg{Rdimtools} & ctb & CRAN & 7367 & Dimension Reduction and Estimation Methods \\
		\cranpkg{remotes} & ctb & CRAN & 3'944'090 & R package installation from remote repositories \\
		\cranpkg{rlang} & ctb & CRAN & 13'269'115 & Functions for base types and core R and tidyverse features \\
		\githubpkg{aertslab}{SCope} & ctb & Github & \notavailable & Visualization of high dimensional single cell data \\
		\biocpkg{slingshot} & ctb & Bioc & 12'085 & Tools for ordering single-cell sequencing \\
		\biocpkg{splatter} & ctb & Bioc & 5015 & Simple simulation of single-cell RNA sequencing data \\
		\cranpkg{tidyverse} & ctb & CRAN &  & Easily install and load packages from the tidyverse \\
		\githubpkg{farrelja}{URD} & ctb & Github & \notavailable & Reconstructing branching trajectories from single-cell RNAseq data \\
		\githubpkg{ManuSetty}{wishbone} & ctb & Github & \notavailable & Identify bifurcating developmental trajectories from single-cell data \\\hline
	\end{tabularx}
\end{table}

>>>>

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{Curriculum Vitae}
\section{Personalia}
\begin{tabular}{ll}
	Name & Robrecht Cannoodt \\
	Date of Birth & January 25th, 1990 \\
	Place of birth & Ghent \\
	Nationality & Belgian \\
	Address & Kerkstraat 75, 9070 Destelbergen \\
	Email & \href{mailto:robrecht.cannoodt@gmail.com}{robrecht.cannoodt@gmail.com} \\
	Webpage & \url{https://www.cannoodt.dev}
\end{tabular}

\section{Professional Experience}
\textit{Ph.D. student in Computer Science} \\
\hspace*{2 ex} Ghent University, Belgium, September 2013 -- now \\
\textit{Laboratory assistant (summer job)} \\
\hspace*{2 ex} Bloedtransfusiecentrum Gent, Belgium, July (2009, 2010, 2011, 2012)

@@

\begin{table}[ht!]
	\caption{\textbf{Contibutions to open-source software.} Following abbreviations denote the relation with respect to the package: \textit{aut} Author, \textit{ctb} Contributor. Yearly download statistics are based on the number of downloads between 2019-10-01 and 2019-11-28. CRAN download statistics are retrieved from the Rstudio CRAN mirror only; other CRAN mirrors do not track download statistics. In addition, many of the dynverse packages have only recently been published on CRAN. For Github repositories, no download statistics could be retrieved. } \label{tab:packages}

	\centering\fontsize{7}{9}\selectfont
	\begin{tabularx}{\linewidth}{|p{2cm}llp{1.5cm}X|}
		\hline
		Name & Role & Host & Downloads per year & Description \\ \hline\hline
		\cranpkg{babelwhale} & aut & CRAN & 6110 & Interacting with Docker and Singularity containers \\
		\cranpkg{diffusionMap} & aut & CRAN & 30'123 & Implements diffusion map method of data parameterization \\
		\githubpkg{dynverse}{dynbenchmark} & aut & Github & \notavailable & Pipeline for benchmarking trajectory inference methods \\
		\cranpkg{dyndimred} & aut & CRAN & 5116 & Applying dimensionality reduction methods \\
		\githubpkg{dynverse}{dyneval} & aut & Github & \notavailable & Evaluating trajectory inference methods \\
		\githubpkg{dynverse}{dynfeature} & aut & Github & \notavailable & Calculating feature importance scores from trajectories \\
		\githubpkg{dynverse}{dyngen} & aut & Github & \notavailable & Simulating single-cell data using gene regulatory networks \\
		\githubpkg{dynverse}{dynguidelines} & aut & Github & \notavailable & User guidelines for trajectory inference \\
		\githubpkg{dynverse}{dynmethods} & aut & Github & \notavailable & A collection of wrappers for trajectory inference methods \\
		\githubpkg{dynverse}{dyno} & aut & Github & \notavailable & A pipeline for inferring, visualising and interpreting trajectories \\
		\cranpkg{dynparam} & aut & CRAN & 3816 & Creating meta-information for parameters \\
		\githubpkg{dynverse}{dynplot} & aut & Github & \notavailable & A simple visualisation library for trajectories \\
		\githubpkg{dynverse}{dynplot2} & aut & Github & \notavailable & A fully customisable visualisation library for trajectories \\
		\githubpkg{dynverse}{dyntoy} & aut & Github & \notavailable & Generating simple toy data of cellular differentiation \\
		\cranpkg{dynutils} & aut & CRAN & 16'999 & Common functionality for the dynverse packages \\
		\cranpkg{dynwrap} & aut & CRAN & 4009 & A common format for trajectories \\
		\cranpkg{GillespieSSA} & aut & CRAN & 7763 & Gillespie's Stochastic Simulation Algorithm (SSA) \\
		\cranpkg{GillespieSSA2} & aut & CRAN & 4181 & Gillespie's Stochastic Simulation Algorithm for Impatient People \\
		\githubpkg{dynverse}{gng} & aut & Github & \notavailable & Growing Neural Gas implemented in Rcpp \\
		\cranpkg{incgraph} & aut & CRAN & 3570 & Incremental graphlet counting for network optimisation \\
		\cranpkg{lmds} & aut & CRAN & 1742 & Landmark Multi-Dimensional Scaling \\
		\cranpkg{princurve} & aut & CRAN & 28'869 & Fits a principal curve in arbitrary dimension \\
		\cranpkg{proxyC} & aut & CRAN & 122'858 & Computes proximity in large sparse matrices \\
		\cranpkg{qsub} & aut & CRAN & 3622 & Running commands remotely on gridengine clusters \\
		\cranpkg{SCORPIUS} & aut & CRAN & 4285 & Inferring developmental chronologies from single-cell RNA sequencing data \\ \hline\hline

		\cranpkg{badger} & ctb & CRAN & 6472 & Query information and generate badge for using in README \\
		\biocpkgl{ClusterSignificance}{Clus\-ter\-Sig\-nif\-i\-cance} &  & Bioc & 935 & Assess if class clusters in dimensionality reduced data representations have a separation different from permuted data \\
		\cranpkg{devtools} & ctb & CRAN & 5'918'700 & Tools to make developing R packages easier \\
		\cranpkg{ggrepel} & ctb & CRAN & 2'018'030 & Repel overlapping text labels away from each other \\
		\githubpkg{soedinglab}{merlot} & ctb & Github & \notavailable & Reconstructing lineage-tree topologies from scRNA-seq data \\
		\cranpkg{mlr} & ctb & CRAN & 176'330 & Machine Learning in R \\
		\biocpkg{monocle} & ctb & Bioc & 34'360 & Clustering, differential expression, and trajectory analysis for single-cell RNA-Seq \\
		\cranpkg{ParamHelpers} & ctb & CRAN & 150'775 & Helpers for Parameters in Black-Box Optimization, Tuning and Machine Learning \\
		\githubpkg{kieranrcampbell}{pseudogp} & ctb & Github & \notavailable & Probabilistic pseudotime for single-cell RNA-seq \\
		\cranpkg{ranger} & ctb & CRAN & 413'641 & A Fast Implementation of Random Forests \\
		\cranpkg{Rdimtools} & ctb & CRAN & 7367 & Dimension Reduction and Estimation Methods \\
		\cranpkg{remotes} & ctb & CRAN & 3'944'090 & R package installation from remote repositories \\
		\cranpkg{rlang} & ctb & CRAN & 13'269'115 & Functions for base types and core R and tidyverse features \\
		\githubpkg{aertslab}{SCope} & ctb & Github & \notavailable & Visualization of high dimensional single cell data \\
		\cranpkg{shadowtext} & ctb & CRAN & 6822 & shadow text for grid and ggplot2 \\
		\biocpkg{slingshot} & ctb & Bioc & 12'085 & Tools for ordering single-cell sequencing \\
		\biocpkg{splatter} & ctb & Bioc & 5015 & Simple simulation of single-cell RNA sequencing data \\
		\cranpkg{tidyverse} & ctb & CRAN & 5'079'398 & Easily install and load packages from the tidyverse \\
		\githubpkg{farrelja}{URD} & ctb & Github & \notavailable & Reconstructing branching trajectories from single-cell RNAseq data \\
		\githubpkg{ManuSetty}{wishbone} & ctb & Github & \notavailable & Identify bifurcating developmental trajectories from single-cell data \\\hline
	\end{tabularx}
\end{table}
